---
title: "Cerrado Land Cover Input"
author: "James D.A. Millington"
date: "July 2022"
output: github_document
#  html_document: 
#    keep_md: yes
#    code_folding: hide
#    df_print: paged
---

```{r}
library(terra)
```

This document presents a record of analysis to create land cover inputs to an application of CRAFTY to Cerrado biome in four states (GO, MT, MS, MG). 

[MapBiomas](https://mapbiomas.org/en) Collection 6 Land Cover raster (.tif) files were downloaded for 2001-2020 for the Cerrado biome via the Toolkits in Google Earth Engine. Spatial resolution was changed from from 30m to 1000m in the run dialogue. Outputs had naming convention _mapbiomas-brazil-collection-60-cerrado-YYYY-1km.tif_ where _YYYY_ is four digit year. 

Combine these raster files into a single multi-layer raster and set `0` values to `No Data`
```{r}
r <- rast("data/mapbiomas6/mapbiomas-brazil-collection-60-cerrado-2001-1km.tif")
for(i in 2002:2020){
  r1 <- rast(paste0("data/mapbiomas6/mapbiomas-brazil-collection-60-cerrado-",i,"-1km.tif"))
  r <- c(r, r1)
}
r[r==0] <- NA
r
plot(r[[1]])
```

This multi-layer raster needs to be clipped to the extent of our four states of interest.

We do this using state boundary vector (.shp) data  downloaded from the [MapBiomas Platform](https://plataforma.brasil.mapbiomas.org/) (Source: IBGE; Year: 2018). 
```{r}
s <- vect("data/vector/dashboard_states-static-layer/dashboard_states-static-layer.shp")
plot(s)
head(s)
```

We can see state ids and state names. States ids for our states are `18369`, `18396`, `18397`, `18399`

The accents on some state names are not being handled well (giving errors/warnings). We'll fix this below before writing out our subset shapefile containing only the four states we want.
```{r}
#state ids in id column
sids <- c(18369, 18396, 18397, 18399)

#subset
G3MGs <- s[s$id %in% sids,]
plot(G3MGs)

head(G3MGs)
#fix GOIAS name
G3MGs[1,'name'] <- "GOIAS"
head(G3MGs)

#write to file
outfile <- "data/vector/G3MGs.shp"
writeVector(G3MGs, outfile, overwrite=TRUE)
```

Read in our subset shapefile and re-project to the CRS of our raster data ready for clipping. 
```{r}
s4<- vect("data/vector/G3MGs.shp")
crsr <- crs(r)
s4 <- terra::project(s4, crsr)
```

Plot one raster layer with the re-projected vector data to check 
```{r}
plot(r[[1]])
plot(s4, add=T)
```

Now clip (crop, mask, trim) the raster data by the re-projected four-state vector (plot one layer to check)
```{r}
rs4 <- crop(r , ext(s4)) 
rs4 <- mask(rs4, s4)
rs4 <- trim(rs4)
rs4
plot(rs4[[1]])
```

Write multi-layer raster to .tif file
```{r}
writeRaster(rs4, "data/mapbiomas6/mapbiomas6-cerrado-G3MGs-2001-2020-1km.tif", overwrite=TRUE)
```

Re-load and plot to check this all worked. 
```{r}
new <- rast("data/mapbiomas6/mapbiomas6-cerrado-G3MGs-2001-2020-1km.tif")
new
plot(new)
```