---
title: "Cerrado Land Cover Input"
author: "James D.A. Millington"
date: "April 2023"
output: github_document
#  html_document: 
#    keep_md: yes
#    code_folding: hide
#    df_print: paged
---

```{r}
library(terra)
library(readxl)
```

This document presents a record of analysis to create land cover inputs to an application of CRAFTY to Cerrado biome in four states (GO, MT, MS, MG). 

[MapBiomas](https://mapbiomas.org/en) Collection 7 Land Cover raster (.tif) files were downloaded for 2001-2021 for the Cerrado biome via the Toolkits in Google Earth Engine. Spatial resolution was changed from from 30m to 1000m in the run dialogue. Outputs had naming convention _mapbiomas-brazil-collection-70-cerrado-YYYY-1km.tif_ where _YYYY_ is four digit year. 

Because we can download mapbiomas data at our desired resolution (whereas previously only 30m was available), we will take a different approach to develop the input region file for CRAFTY [than used for CRAFTY Brazil](https://github.com/jamesdamillington/BrazilInputMaps). Rather than creating the BaseRaster first from a rasterised vector, we'll use the mapbiomas raster as a base and work from there. 

So, first. combine the mapbiomas raster files into a single multi-layer raster and set `0` values to `No Data`
```{r}
r <- rast("data/raster/mapbiomas7/mapbiomas-brazil-collection-70-cerrado-2001-1km.tif")
for(i in 2002:2021){
  r1 <- rast(paste0("data/raster/mapbiomas7/mapbiomas-brazil-collection-70-cerrado-",i,"-1km.tif"))
  r <- c(r, r1)
}
r[r==0] <- NA
r
plot(r[[1]])
```

This multi-layer raster needs to be clipped to the extent of our four states of interest.

We do this using municiaplity and state boundary vector (.shp) data 
```{r}
BRadmin <- vect("data/vector/BRadmin.shp")
plot(BRadmin)
head(BRadmin)
```

We can see state ids and state names. States ids for our states are 50, 51, 52, 31



The accents on some state names are not being handled well (giving errors/warnings). We'll fix this below before writing out our subset shapefile containing only the four states we want.
```{r}
#state ids in id column
sids <- c(50, 51, 52, 31)

#subset
G3MGsadmin <- BRadmin[BRadmin$State %in% sids,]
plot(G3MGsadmin)

head(G3MGsadmin)

#write to file
outfile <- "data/vector/G3MGsadmin.shp"
writeVector(G3MGsadmin, outfile, overwrite=TRUE)
```

Read in our subset shapefile and re-project to the CRS of our raster data ready for clipping. 
```{r}
s4<- vect("data/vector/G3MGsadmin.shp")
crsr <- crs(r)
s4 <- terra::project(s4, crsr)
```

Plot one raster layer with the re-projected vector data to check 
```{r}
plot(r[[1]])
plot(s4, add=T)
```

Now clip (crop, mask, trim) the raster data by the re-projected four-state vector (plot one layer to check)
```{r}
rs4 <- crop(r , ext(s4)) 
rs4 <- mask(rs4, s4)
rs4 <- trim(rs4)
rs4
plot(rs4[[1]])
plot(s4, add=T)
```

Write multi-layer raster to .tif file
```{r}
writeRaster(rs4, "data/raster/mapbiomas7/mapbiomas7-cerrado-G3MGs-2001-2021-1km.tif", overwrite=TRUE)
```

Re-load and plot to check this all worked. 
```{r}
new <- rast("data/raster/mapbiomas7/mapbiomas7-cerrado-G3MGs-2001-2021-1km.tif")
new
plot(new)

#check classes present in each year
unique(new, incomparables = TRUE)

```

Now reclassify (no need for disaggregation like for CRAFTY-Brazil, because collection 7 include classes for soybeans etc)

```{r}
classification <- read_excel(paste0("data/tables/mapbiomas7/MapBiomas_CRAFTY_classifications_v7.xlsx"), sheet = 'mb7-reclass1', range="B1:C37", col_names=T)  
map <- classify(new, rcl=as.matrix(classification))                 #classify
plot(map)
writeRaster(map, "data/raster/mapbiomas7/mapbiomas7-cerrado-G3MGs-2001-2021-1km-reclass1.tif", overwrite=T) #write!
```

Create raster maps of states and municipalities (for region.csv), usnf dimensions etc from 'new' `rast`

```{r}
G3MGsadmin_latlon <- rast(
  nrows=nrow(new),
  ncols=ncol(new),
  nlyrs=1,
  xmin=xmin(new),
  xmax=xmax(new),
  ymin=ymin(new),
  ymax=ymax(new),
  crs=crs(new),
  resolution=res(new),
  vals=NA
)

G3MGsmunis_r_latlon <- rasterize(x=s4, y=G3MGsadmin_latlon, field = "CD_GEOCMUn")
G3MGsstates_r_latlon <- rasterize(x=s4, y=G3MGsadmin_latlon, field = "State")

G3MGsmunis_r_latlon <- mask(G3MGsmunis_r_latlon, new[[1]])
G3MGsstates_r_latlon <- mask(G3MGsstates_r_latlon, new[[1]])

plot(G3MGsmunis_r_latlon)
plot(G3MGsstates_r_latlon)

writeRaster(G3MGsmunis_r_latlon, "data/raster/socecon/G3MGsmunis_r_latlon.tif", overwrite=TRUE)
writeRaster(G3MGsstates_r_latlon, "data/raster/socecon/G3MGsstates_r_latlon.tif", overwrite=TRUE)
```

